A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-05-22 paper/paper1/ms.tex
Add paragraph stating the layout of the paper at end of intro

\begin{document}
\shorttitle{HPF spectral analysis}
\shortauthors{TBD}

@@

Uncertainty in system ages, evaporation timescales, and dominating physical mechanisms, degrade our ability to foretell if any given planet will exhibit ongoing signatures of atmospheric escape.  Episodic stellar wind gusts and other forms of astrophysical variance could also subdue the appearance of atmospheric escape, even where we expect it most.  The 48 published non-detections of atmospheric escape \citep{2022arXiv221116243D} must encode these natural whims in a way that we have not yet disentangled. Nevertheless, we can boost our chances of witnessing active and significant atmospheric escape by targeting sources that seem predisposed to loss.  These intrinsic or extrinsic factors may include proximity to the host star, low surface gravity, and high energy incident radiation.  The small planet radius gap between 1.5$-$2.0 Earth radius range could be expected to harbor atmospheres that have not yet shed their gaseous envelopes, still in the act of eroding to their final rocky state.

Inflated hot Saturns stand out as an especially extreme category that should exhibit mass loss.  Their low gravitational potentials should let go of their atmospheres more readily than their steadfast Hot Jupiter counterparts.  Their lower gravity yields larger atmospheric scale heights making them easier to detect in transmission spectroscopy.  And their large transit depths and short periods should make them easy to detect in large numbers, like Hot Jupiters.

@@

Alternatively---and most consequentially for atmospheric escape---another explanation may prevail.  Inflated hot Saturns may either form \emph{in-situ} or effectively migrate to close-in orbital separations \citep{2018ARA&A..56..175D}, but once they arrive, the intense irradiation overheats the planet.  This heating leads to runaway inflation and ultimately complete disintegration.  In this scenario, the inflationary half-life may be so short as to lead to a low probability of observing members in the class among the overall population, and hence the apparent lack of inflated sub-Saturns \citep{2023ApJ...945L..36T}.

\citet{2011ApJ...738....1B} predicted runaway inflation of hot Saturns as a consequence of the Ohmic dissipation mechanism.  Here, lightly thermally ionized atmospheric flows induce drag in a planetary magnetic field, weakly coupling the stellar incident energy into the planetary interior.  \citet{2018AJ....155..214T} recognized the above two hypotheses for the lack of sub-Saturns, while statistically favoring Ohmic dissipation as the mechanism responsible for inflating Hot Jupiters.  Therefore Ohmic dissipation stands out as a leading physical mechanism for inflation and mass loss.

@@

\section{Observations}
\subsection{Habitable Zone Planet Finder (HPF)}

@@

\section{Analysis}

@@

\subsection{HPF analysis I: RV fitting}

@@

\subsection{HPF analysis II: \ion{He}{1} 10833 \AA}
The preparation for analyzing the Helium feature followed a similar but slightly different procedure than that used for the RV, namely, we used \texttt{Goldilocks}\footnote{\url{https://github.com/grzeimann/Goldilocks_Documentation}} for 2D echellogram reduction.  This tool outputs 1D extracted spectra for the target, and two reference fibers: blank sky and a laser frequency comb \citep[LFC,][]{2019Optic...6..233M}.  The observations were usually acquired with the LFC turned off, so this unilluminated spectrum was discarded.

@@

\section{Atmospheric Escape}

@@

\section{Causal Interpretations} \label{secDiscuss}

@@

\subsection{3D Hydrodynamic simulations}
\authorcomment1{Insert Morgan Macleod text here.}
\textcolor{lightgray}{Habitasse platea dictumst vestibulum rhoncus est pellentesque elit ullamcorper dignissim. Odio ut sem nulla pharetra. Odio ut sem nulla pharetra diam sit amet nisl suscipit. Accumsan tortor posuere ac ut. Ornare arcu dui vivamus arcu felis bibendum ut tristique. Nunc faucibus a pellentesque sit amet porttitor eget. Mauris rhoncus aenean vel elit. Sed vulputate mi sit amet mauris commodo. Quis lectus nulla at volutpat diam ut venenatis. Sed id semper risus in hendrerit gravida rutrum quisque non. Est placerat in egestas erat imperdiet sed. Viverra adipiscing at in tellus integer feugiat scelerisque varius. Quam viverra orci sagittis eu volutpat odio facilisis mauris sit. Morbi tristique senectus et netus et malesuada. Duis tristique sollicitudin nibh sit amet. Libero nunc consequat interdum varius sit amet mattis vulputate. Velit ut tortor pretium viverra suspendisse potenti nullam ac. Viverra accumsan in nisl nisi scelerisque eu ultrices vitae. Quis commodo odio aenean sed adipiscing diam donec.}

@@

\begin{deluxetable}{lLLCR}
    \tablewidth{0pc}
    \tabletypesize{\scriptsize}
    \tablecaption{
        Coarse Inflation Timescales for Hot Jupiters
        \label{tabHeSM}
    }
    \tablehead{
        \colhead{Planet}   &
        \colhead{Mass} &
        \colhead{Radius} &
        \colhead{$T_\mathrm{eq}$} &
        \colhead{$\tau_\mathrm{infl}$} \\
        \colhead{}   &
        \colhead{$M_\mathrm{Jup}$} &
        \colhead{$R_\mathrm{Jup}$}   &
        \colhead{K} &
        \colhead{Myr}
    }
    \startdata
    \input{tables/tab_HeSM.tex}
    \enddata
    %\tablecomments{Inflation timescale assumes $\epsilon=0.01$.}  Not true, we use the empirical estimate
\end{deluxetable}

\section{Discussion}

>>>>

\begin{document}
\shorttitle{Runaway Mass Loss in HAT-P-67b}
\shortauthors{TBD}

@@

Uncertainty in system ages, evaporation timescales, and dominating physical mechanisms, degrade our ability to foretell if any given planet will exhibit ongoing signatures of atmospheric escape.  Episodic stellar wind gusts and other forms of astrophysical variance could also subdue the appearance of atmospheric escape, even where we expect it most.  The 48 published non-detections of atmospheric escape \citep{2022arXiv221116243D} must encode these natural whims in a way that we have not yet disentangled. Nevertheless, we can boost our chances of witnessing active and significant atmospheric escape by targeting sources that seem predisposed to loss.  These intrinsic or extrinsic factors may include proximity to the host star, low surface gravity, and high energy incident radiation.

Inflated hot Saturns ($T_\mathrm{eq}>1000\;$K, $M_p \lesssim 0.3 M_\mathrm{Jup}$) stand out as an especially extreme category that should exhibit mass loss.  Their low gravitational potentials should let go of their atmospheres more readily than their steadfast hot Jupiter counterparts.  Their lower gravity yields larger atmospheric scale heights making them easier to detect in transmission spectroscopy.  And their large transit depths and short periods should make them easy to detect in large numbers, like hot Jupiters.

@@

Alternatively---and most consequentially for atmospheric escape---another explanation may prevail.  Inflated hot Saturns may either form \emph{in-situ} or effectively migrate to close-in orbital separations \citep{2018ARA&A..56..175D}, but once they arrive, the intense irradiation overheats the planet.  This heating leads to runaway inflation and ultimately complete disintegration.  In this scenario, the inflationary half-life becomes so short that the probability of observing members in the class decreases sharply with density, causing the apparent lack of inflated sub-Saturns \citep{2023ApJ...945L..36T}.

\citet{2011ApJ...738....1B} predicted runaway inflation of hot Saturns as a consequence of the Ohmic dissipation mechanism.  Here, lightly thermally ionized atmospheric flows induce drag in a planetary magnetic field, weakly coupling the stellar incident energy into the planetary interior.  \citet{2018AJ....155..214T} recognized the above two hypotheses for the lack of sub-Saturns, while statistically favoring Ohmic dissipation as the mechanism responsible for inflating hot Jupiters.  Therefore Ohmic dissipation stands out as a leading physical mechanism for inflation and mass loss.

@@

We assemble both archival, previously published, and new observations of the HAT-P-67 system, chronicled in Section \ref{secObs}.  In Section \ref{secAnalysis} we refine the stellar and planet properties based on those observations, including distance, radius, and orbit re-analyses.  The Helium excess detection is presented in \ref{secHeAnalysis}, with deep dives on the signal's trend with orbital phase.  We use this orbital structure to reconstruct the geometry and mass loss of the escaping Helium exosphere, with both 1D and 3D models (Section \ref{secResults}).  We assess the physical mechanisms (Section \ref{secPhysicalMechanisms}) giving rise to the escaping atmosphere, weighing the distinctive predictions of XUV irradation and Ohmic dissipation.  Finally in Section \ref{secDiscuss}, we question the assumptions in our approach, discuss the overall congruence of predictions and observations, and highlight some implications for future exosphere studies.

\section{Observations}\label{secObs}
\subsection{Habitable Zone Planet Finder (HPF)}

@@

\section{Analysis} \label{secAnalysis}

@@

\subsection{HPF analysis I: RV fitting}\label{secRVfit}

@@

\subsection{HPF analysis II: \ion{He}{1} 10833 \AA} \label{secHeAnalysis}
The preparation for analyzing the Helium feature followed a similar but slightly different procedure than that used for the RV, namely, we used \texttt{Goldilocks}\footnote{\url{https://github.com/grzeimann/Goldilocks_Documentation}} for 2D echellogram reduction.  This tool outputs 1D extracted spectra for the target, and two reference fibers: blank sky and a laser frequency comb \citep[LFC,][]{2019Optic...6..233M}.  The observations were usually acquired with the LFC turned off, so this unilluminated spectrum was discarded.

@@

\section{Atmospheric Escape}\label{secResults}

@@

\section{Causal Interpretations} \label{secPhysicalMechanisms}

@@

\subsection{3D Hydrodynamic simulations}\label{sec3Dsimulations}
\authorcomment1{Insert Morgan Macleod text here.}
\textcolor{lightgray}{Habitasse platea dictumst vestibulum rhoncus est pellentesque elit ullamcorper dignissim. Odio ut sem nulla pharetra. Odio ut sem nulla pharetra diam sit amet nisl suscipit. Accumsan tortor posuere ac ut. Ornare arcu dui vivamus arcu felis bibendum ut tristique. Nunc faucibus a pellentesque sit amet porttitor eget. Mauris rhoncus aenean vel elit. Sed vulputate mi sit amet mauris commodo. Quis lectus nulla at volutpat diam ut venenatis. Sed id semper risus in hendrerit gravida rutrum quisque non. Est placerat in egestas erat imperdiet sed. Viverra adipiscing at in tellus integer feugiat scelerisque varius. Quam viverra orci sagittis eu volutpat odio facilisis mauris sit. Morbi tristique senectus et netus et malesuada. Duis tristique sollicitudin nibh sit amet. Libero nunc consequat interdum varius sit amet mattis vulputate. Velit ut tortor pretium viverra suspendisse potenti nullam ac. Viverra accumsan in nisl nisi scelerisque eu ultrices vitae. Quis commodo odio aenean sed adipiscing diam donec.}

@@

\begin{deluxetable}{lLLCR}
    \tablewidth{0pc}
    \tabletypesize{\scriptsize}
    \tablecaption{
        Coarse Inflation Timescales for hot Jupiters
        \label{tabHeSM}
    }
    \tablehead{
        \colhead{Planet}   &
        \colhead{Mass} &
        \colhead{Radius} &
        \colhead{$T_\mathrm{eq}$} &
        \colhead{$\tau_\mathrm{infl}$} \\
        \colhead{}   &
        \colhead{$M_\mathrm{Jup}$} &
        \colhead{$R_\mathrm{Jup}$}   &
        \colhead{K} &
        \colhead{Myr}
    }
    \startdata
    \input{tables/tab_HeSM.tex}
    \enddata
    %\tablecomments{Inflation timescale assumes $\epsilon=0.01$.}  Not true, we use the empirical estimate
\end{deluxetable}

\section{Discussion}\label{secDiscuss}
